Blame and Coercion: Together Again for the Second Time

### 2019-10-17 paper.tex
minor
give LazyD C a name: C_D

\newcommand{\judgeDCreduce}[2]{#1 \longrightarrow_{\mathcal{D}} #2}
\newcommand{\judgeDCreduceTrans}[2]{#1 \longrightarrow_{\mathcal{D}}^{*} #2}
\newcommand{\judgeDCeval}[2]{#1 \Downarrow_{\mathcal{D}} #2}
\newcommand{\redrule}[3]{#1 & \longrightarrow_\mathcal{D} & #2 & #3\\}

@@

	Reduction \fbox{$ \judgeDCreduce{s}{s} $}
	\[
	\begin{array}{rclr}
	\redrule{
		\sOOinspect{\eOOvar{x}}{\rho}{\kappa}
	}{	
		\sOOreturn{\rho[x]}{\kappa}
	}{}
	\redrule{
		\sOOinspect{\eOOsole}{\rho}{\kappa}
	}{
		\sOOreturn{\vOOtt}{\kappa}
	}{}
	\redrule{
		\sOOinspect{\eOOlam{T_1}{T_2}{x}{e}}{\rho}{\kappa}
	}{
		\sOOreturn{(\vOOfun{\rho}{x}{e})}{\kappa}
	}{}
	\redrule{
		\sOOinspect{(\eOOcons{e_1}{e_2})}{\rho}{\kappa}
	}{
		\sOOinspect{e_1}{E}{}
	}{}
	\redrule{
		\sOOinspect{(\eOOinl{e})}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{(\kOOinl{\kappa})}
	}{}
	\redrule{
	\sOOinspect{(\eOOinr{e})}{\rho}{\kappa}
	}{
	\sOOinspect{e}{\rho}{(\kOOinr{\kappa})}
	}{}
	\redrule{
		\sOOinspect{(\eOOapp{e_1}{e_2})}{\rho}{\kappa}
	}{
\sOOinspect{e_1}{\rho}{(\kOOappI{E}{e_2}{\kappa})}}{}
\redrule{
\sOOinspect{(\eOOcar{e})}{\rho}{\kappa}}{
\sOOinspect{e}{\rho}{(\kOOcar{\kappa})}}{}
\redrule{
	\sOOinspect{(\eOOcdr{e})}{\rho}{\kappa}}{
	\sOOinspect{e}{\rho}{(\kOOcdr{\kappa})}}{}
\redrule{
\sOOinspect{(\eOOcase{e_1}{e_2}{e_3})}{\rho}{\kappa}}{
\sOOinspect{e_1}{\rho}{(\kOOcaseI{E}{e_2}{e_3}{\kappa})}}{}
	\redrule{
		\sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{\langle\cOOcast{T_1}{l}{T_2}\rangle\kappa}
	}{}

@@

	Evaluation \fbox{$\judgeDCeval{e}{o}$}
	\[
	\inference{
		\sOOinspect{e}{\emptyset}{cont(\hckOOmt)} \longrightarrow_{B}^{*} 
		\sOOhalt{o}
	}{
		\judgeDCeval{e}{o}
	}
	\]

	\caption{dynamic semantics of the cast calculi as a CEK machine}
	\label{machine-cekc}
\end{figure}

@@

The reduction relation $ \judgeDCreduce{s}{s} $ depends on $ applyCast $, a 
function depending on blame strategies. We will give the definitions of $ 
applyCast $ shortly.
To evaluate a cast 
expression, the machine move the 
cast to the continuation and evaluate the inner expression. To apply a casted 
function, the machine firstly cast $ v_1 $, the operand, then apply the casted 
operand to $ v_2 $, the un-casted function, and finally cast the result of the 
function application. To take out the first (resp. second) part of a casted 
pair, the machine firstly take out the first (resp. second) part of $ v $, the 
un-casted pair, and cast the result. To case-split a sum, the machine looks 
deep inside the value to find out whether it is a left injection or a right 
one.
\todo[inline]{Change the rules for case to be small step, i.e.,
	the case function should not be recursive. -JS}
Along the way, the machine accumulates the pending casts onto the 
continuations. To cast a value, the machine try to apply the cast to the value. 
If the cast succeeds, the machine returns the result to the next continuation. 
If the cast fails, the machine halts with the blame label.

The reflexive transitive closure of reduction ($ \judgeDCreduceTrans{s}{s} $) 
and evaluation ($ \judgeDCeval{e}{o} $) are standard.

@@

\figref{fig:applyCast-D-C} defines the $ applyCast $ for \lazyD. 
We conjecture that with this $ applyCast $, the CEK machine in 
\figref{machine-cekc} would agree with the \lazyD{} cast calculus in 
\citet{siek2009exploring}.

@@

We use bisimulation to proof equivalence between the two machines.
\todo[inline]{Which two machines? -JS}
The bisimulation relation of states ($\eqvDS{C}{s}{s}$) is mostly derived from 
the definitions of $ \mathcal{D} $. Exceptions include the relations 
between continuations, between values, and from continuations to 
pre-continuations. We omit subscripts when there is no ambiguity.

@@

\begin{lemma}[Weak Bisimulation between $ \mathcal{S}(\cdot) $ and $ 
	\mathcal{D} $]
	\label{thm:surely-monoidic-reduce}
	For all cast representations $ C $, 
	if $ C $ is surely \lazyD{} and monoidic,
	$ s_1 \in \mathcal{D} $,
	$ s_2 \in \mathcal{S}(C) $, and
	$ s_1 \approx s_2 $, 
	\begin{enumerate}
		\item If $ \judgeDCreduce{s_1}{s_3} $,
		then
		there exist $ s_5 \in \mathcal{D} $, $ s_4 \in \mathcal{S}(C) $,
		such that \begin{itemize}
			\item $ s_5 \approx s_4 $, and
			\item $ \judgeSreduce{C}{s_2}{s_4} $,
			\item $ \judgeDCreduceTrans{s_3}{s_5} $
		\end{itemize}
		\item If $ \judgeSreduce{C}{s_2}{s_4} $,
		then
		there exist $ s_3 \in \mathcal{D} $,
		such that \begin{itemize}
			\item $ s_3 \approx s_4 $
			\item $ \judgeDCreduceTrans{s_1}{s_3} $
		\end{itemize}
	\end{enumerate}
\end{lemma}
\begin{proof}
	See the supplementary material.
\end{proof}

\begin{lemma}[Surely \lazyD\ and Monoidic Cast Representations Respect 
	$ \mathcal{D} $]
	\label{thm:surely-monoidic-eval}
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ and $ C $ is surely 
	\lazyD{} and monoidic,
	\[
	\judgeDCeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{C}{e}{o}
	\]
\end{lemma}
\begin{proof}
  Immediately from Lemma~\ref{thm:surely-monoidic-reduce}.
  \todo[inline]{I'm guessing that you prove this lemma by induction,
    using Lemma~\ref{thm:surely-monoidic-reduce}.
    Proofs by induction are not ``immediate''. -JS}
\end{proof}

@@

\begin{theorem}[\lazyD{} hypercoercion Respect $ \mathcal{D} $]
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ 
	\[
	\judgeDCeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{H}{e}{o}
	\]
\end{theorem}
\begin{proof}
	Immediately from Lemma~\ref{thm:surely-monoidic-eval},
	Lemma~\ref{thm:hc-surely-lazyD}, and 
	Proposition~\ref{thm:hc-monoid}.
\end{proof}

@@

\begin{theorem}[Surely \lazyD\ Cast Representations Respect 
	$ \mathcal{D} $]
	\label{thm:surely-lazyD-correct}
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ and $ C $ is surely 
	\lazyD{}
	\[
	\judgeDCeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{C}{e}{o}
	\]
\end{theorem}
\begin{proof}
	Immediately from Lemma~\ref{thm:surely-monoidic-eval} and 
	Proposition~\ref{thm:surely-lazyD-eqv}.
\end{proof}

>>>>

\newcommand{\judgeCreduce}[2]{#1 \longrightarrow_{\mathcal{C}} #2}
\newcommand{\judgeCreduceTrans}[2]{#1 \longrightarrow_{\mathcal{C}}^{*} #2}
\newcommand{\judgeCeval}[2]{#1 \Downarrow_{\mathcal{C}} #2}
\newcommand{\redrule}[3]{#1 & \longrightarrow_\mathcal{C} & #2 & #3\\}

@@

	Reduction \fbox{$ \judgeCreduce{s}{s} $}
	\[
	\begin{array}{rclr}
	\redrule{
		\sOOinspect{\eOOvar{x}}{\rho}{\kappa}
	}{	
		\sOOreturn{\rho[x]}{\kappa}
	}{}
	\redrule{
		\sOOinspect{\eOOsole}{\rho}{\kappa}
	}{
		\sOOreturn{\vOOtt}{\kappa}
	}{}
	\redrule{
		\sOOinspect{\eOOlam{T_1}{T_2}{x}{e}}{\rho}{\kappa}
	}{
		\sOOreturn{(\vOOfun{\rho}{x}{e})}{\kappa}
	}{}
	\redrule{
		\sOOinspect{(\eOOcons{e_1}{e_2})}{\rho}{\kappa}
	}{
		\sOOinspect{e_1}{E}{}
	}{}
	\redrule{
		\sOOinspect{(\eOOinl{e})}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{(\kOOinl{\kappa})}
	}{}
	\redrule{
	\sOOinspect{(\eOOinr{e})}{\rho}{\kappa}
	}{
	\sOOinspect{e}{\rho}{(\kOOinr{\kappa})}
	}{}
	\redrule{
		\sOOinspect{(\eOOapp{e_1}{e_2})}{\rho}{\kappa}
	}{
\sOOinspect{e_1}{\rho}{(\kOOappI{E}{e_2}{\kappa})}}{}
\redrule{
\sOOinspect{(\eOOcar{e})}{\rho}{\kappa}}{
\sOOinspect{e}{\rho}{(\kOOcar{\kappa})}}{}
\redrule{
	\sOOinspect{(\eOOcdr{e})}{\rho}{\kappa}}{
	\sOOinspect{e}{\rho}{(\kOOcdr{\kappa})}}{}
\redrule{
\sOOinspect{(\eOOcase{e_1}{e_2}{e_3})}{\rho}{\kappa}}{
\sOOinspect{e_1}{\rho}{(\kOOcaseI{E}{e_2}{e_3}{\kappa})}}{}
	\redrule{
		\sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{\langle\cOOcast{T_1}{l}{T_2}\rangle\kappa}
	}{}

@@

	Evaluation \fbox{$\judgeCeval{e}{o}$}
	\[
	\inference{
		\sOOinspect{e}{\emptyset}{cont(\hckOOmt)} \longrightarrow_{B}^{*} 
		\sOOhalt{o}
	}{
		\judgeCeval{e}{o}
	}
	\]

	\caption{Dynamic semantics of the cast calculi as a CEK machine ($ 
	\mathcal{C} $)}
	\label{machine-cekc}
\end{figure}

@@

The reduction relation $ \judgeCreduce{s}{s} $ depends on $ applyCast $, a 
function depending on blame strategies. We will give the definitions of $ 
applyCast $ shortly.
To evaluate a cast 
expression, the machine move the 
cast to the continuation and evaluate the inner expression. To apply a casted 
function, the machine firstly cast $ v_1 $, the operand, then apply the casted 
operand to $ v_2 $, the un-casted function, and finally cast the result of the 
function application. To take out the first (resp. second) part of a casted 
pair, the machine firstly take out the first (resp. second) part of $ v $, the 
un-casted pair, and cast the result. To case-split a sum, the machine looks 
deep inside the value to find out whether it is a left injection or a right 
one.
\todo[inline]{Change the rules for case to be small step, i.e.,
	the case function should not be recursive. -JS}
Along the way, the machine accumulates the pending casts onto the 
continuations. To cast a value, the machine try to apply the cast to the value. 
If the cast succeeds, the machine returns the result to the next continuation. 
If the cast fails, the machine halts with the blame label.

The reflexive transitive closure of reduction ($ \judgeCreduceTrans{s}{s} $) 
and evaluation ($ \judgeCeval{e}{o} $) are standard.

@@

\figref{fig:applyCast-D-C} defines the $ applyCast $ for \lazyD. 
We conjecture that with this $ applyCast $, the CEK machine in 
\figref{machine-cekc} would agree with the \lazyD{} cast calculus in 
\citet{siek2009exploring}. We denote this \lazyD{} machine by $ 
\mathcal{C}_{\mathtt{D}} $.

@@

We use bisimulation to proof equivalence between the two machines.
\todo[inline]{Which two machines? -JS}
The bisimulation relation of states ($\eqvDS{C}{s}{s}$) is mostly derived from 
the definitions of $ \mathcal{D} $. Exceptions include the relations from 
continuations to pre-continuations, between continuations, and between values. 
We omit subscripts when there is no ambiguity.

@@

\begin{lemma}[Weak Bisimulation between $ \mathcal{S}(\cdot) $ and $ 
	\mathcal{D} $]
	\label{thm:surely-monoidic-reduce}
	For all cast representations $ C $, 
	if $ C $ is surely \lazyD{} and monoidic,
	$ s_1 \in \mathcal{D} $,
	$ s_2 \in \mathcal{S}(C) $, and
	$ s_1 \approx s_2 $, 
	\begin{enumerate}
		\item If $ \judgeCreduce{s_1}{s_3} $,
		then
		there exist $ s_5 \in \mathcal{D} $, $ s_4 \in \mathcal{S}(C) $,
		such that \begin{itemize}
			\item $ s_5 \approx s_4 $, and
			\item $ \judgeSreduce{C}{s_2}{s_4} $,
			\item $ \judgeCreduceTrans{s_3}{s_5} $
		\end{itemize}
		\item If $ \judgeSreduce{C}{s_2}{s_4} $,
		then
		there exist $ s_3 \in \mathcal{D} $,
		such that \begin{itemize}
			\item $ s_3 \approx s_4 $
			\item $ \judgeCreduceTrans{s_1}{s_3} $
		\end{itemize}
	\end{enumerate}
\end{lemma}
\begin{proof}
	See the supplementary material.
\end{proof}

\begin{lemma}[Surely \lazyD\ and Monoidic Cast Representations Respect 
	$ \mathcal{D} $]
	\label{thm:surely-monoidic-eval}
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ and $ C $ is surely 
	\lazyD{} and monoidic,
	\[
	\judgeCeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{C}{e}{o}
	\]
\end{lemma}
\begin{proof}
  Immediately from Lemma~\ref{thm:surely-monoidic-reduce}.
  \todo[inline]{I'm guessing that you prove this lemma by induction,
    using Lemma~\ref{thm:surely-monoidic-reduce}.
    Proofs by induction are not ``immediate''. -JS}
\end{proof}

@@

\begin{theorem}[\lazyD{} hypercoercion Respect $ \mathcal{D} $]
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ 
	\[
	\judgeCeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{H}{e}{o}
	\]
\end{theorem}
\begin{proof}
	Immediately from Lemma~\ref{thm:surely-monoidic-eval},
	Lemma~\ref{thm:hc-surely-lazyD}, and 
	Proposition~\ref{thm:hc-monoid}.
\end{proof}

@@

\begin{theorem}[Surely \lazyD\ Cast Representations Respect 
	$ \mathcal{D} $]
	\label{thm:surely-lazyD-correct}
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ and $ C $ is surely 
	\lazyD{}
	\[
	\judgeCeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{C}{e}{o}
	\]
\end{theorem}
\begin{proof}
	Immediately from Lemma~\ref{thm:surely-monoidic-eval} and 
	Proposition~\ref{thm:surely-lazyD-eqv}.
\end{proof}
